---
title: "Figures and tables for COV002 analysis"
output: html_document
---


#1. Setup

a) Run "1. COV002 VE analysis GitHub" or download generated model fits:

- fit_cov_sens_spec.rds = final model, average population characteristics
- fit_cov_sens_spec_fig.rds = final model, reference population characteristics

```{r}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/GitHub files")
setwd("C:/Users/lrw19/OneDrive - Imperial College London/Documents/PhD Project/RStan Model/GitHub files")
mod_av <- readRDS("fit_cov_sens_spec.rds")#final model, average population characteristics
mod_30 <- readRDS("fit_cov_sens_spec_fig.rds")#final model, reference population characteristics
```


```{r}
library("ggplot2")
library("dplyr")
lci<-function(x){quantile(x, 0.025)}
uci<-function(x){quantile(x, 0.975)}
```


#2. Calculate relative estimates

To get the estimates for the text in the paper (e.g. difference in FOI by age), use the averages model.

```{r absolute parameter ests all}
pars_rel<-c("FOI_diff_cov1", "FOI_diff_cov2", "FOI_diff_cov3",
            "OR_ps_diff_cov1", "OR_ps_diff_cov2", "OR_ps_diff_cov3", "OR_ps_diff_cov4", "OR_ps_diff_cov5",
            "PS_diff_cov1", "PS_diff_cov2", "PS_diff_cov3", "PS_diff_cov4", "PS_diff_cov5",
            "RR_in_diff_cov3", "RR_in_diff_cov3_30yrs", 
            "RR_pr_diff_cov3", "RR_pr_diff_cov5", "RR_pr_diff_cov3_30yrs"
           )
```


```{r convert to dataframe}
dat_mod_fin<-as.data.frame(mod_av)
dat_mod_fin<-dat_mod_fin%>%
  select(c(pars_rel))
dat_mod_fin

summary(dat_mod_fin)
```

```{r estimates}
summary_dat_means<-
  dat_mod_fin%>%
summarise(across(c(FOI_diff_cov1, FOI_diff_cov2, FOI_diff_cov3,
            OR_ps_diff_cov1, OR_ps_diff_cov2, OR_ps_diff_cov3, OR_ps_diff_cov4, OR_ps_diff_cov5,
            PS_diff_cov1, PS_diff_cov2, PS_diff_cov3, PS_diff_cov4, PS_diff_cov5,
            RR_in_diff_cov3, RR_in_diff_cov3_30yrs, 
            RR_pr_diff_cov3, RR_pr_diff_cov3_30yrs, RR_pr_diff_cov5),
             .fns = list(mean=mean))
)

summary_dat_lci<-
  dat_mod_fin%>%
summarise(across(c(FOI_diff_cov1, FOI_diff_cov2, FOI_diff_cov3,
            OR_ps_diff_cov1, OR_ps_diff_cov2, OR_ps_diff_cov3, OR_ps_diff_cov4, OR_ps_diff_cov5,
            PS_diff_cov1, PS_diff_cov2, PS_diff_cov3, PS_diff_cov4, PS_diff_cov5,
            RR_in_diff_cov3, RR_in_diff_cov3_30yrs, 
            RR_pr_diff_cov3, RR_pr_diff_cov3_30yrs, RR_pr_diff_cov5),
             .fns = list(lci=lci))
)

summary_dat_uci<-
  dat_mod_fin%>%
summarise(across(c(FOI_diff_cov1, FOI_diff_cov2, FOI_diff_cov3,
            OR_ps_diff_cov1, OR_ps_diff_cov2, OR_ps_diff_cov3, OR_ps_diff_cov4, OR_ps_diff_cov5,
            PS_diff_cov1, PS_diff_cov2, PS_diff_cov3, PS_diff_cov4, PS_diff_cov5,
            RR_in_diff_cov3, RR_in_diff_cov3_30yrs, 
            RR_pr_diff_cov3, RR_pr_diff_cov3_30yrs, RR_pr_diff_cov5),
             .fns = list(uci=uci))
)
```

```{r calculate differences}
#FOI - age
round(summary_dat_means$FOI_diff_cov3_mean, digits = 2)
round(summary_dat_lci$FOI_diff_cov3_lci, digits = 3)
round(summary_dat_uci$FOI_diff_cov3_uci, digits = 2)

summary_dat_means
summary_dat_lci
summary_dat_uci
```


# 3. Table of average estimates

## a. choose variables

This is run with model with age centered around the mean (~45yrs)

```{r average parameter estimates}
pars_ab_av<-c("FOI_p",
           "PS_p",
           "VE_in",
           "VE_pr",
           "VE_sym",
           "VE_asym"
           )

```

```{r convert to dataframe bias}
dat_mod_av<-as.data.frame(mod_av)
dat_mod_av<-dat_mod_av%>%
  select(c(pars_ab_av))
dat_mod_av
```

## b. create summary data for table

```{r create table bias}
summary_dat_means<-
  dat_mod_av%>%
summarise(across(c(FOI_p, PS_p, VE_in, VE_pr, VE_sym, VE_asym),
             .fns = list(mean=mean))
)
summary_dat_lci<-
  dat_mod_av%>%
summarise(across(c(FOI_p, PS_p, VE_in, VE_pr, VE_sym, VE_asym),
             .fns = list(lci=lci))
)
summary_dat_uci<-
  dat_mod_av%>%
summarise(across(c(FOI_p, PS_p, VE_in, VE_pr, VE_sym, VE_asym),
             .fns = list(uci=uci))
)
params= c(c("FOI", "PS_p", "VE_in", "VE_pr", "VE_sym", "VE_asym"))
covariates=c(pars_ab_av)
means=as.numeric(summary_dat_means[1,])
ucis=as.numeric(summary_dat_uci[1,])
lcis=as.numeric(summary_dat_lci[1,])

summarydat<-data.frame(params, covariates, means, ucis, lcis)
summarydat$covariates <- as.character(summarydat$covariates)
summarydat$covariates <- factor(summarydat$covariates, levels=unique(summarydat$covariates))
summarydat$covs_written<-c("FOI", "PS_p", "VEin", "VEpr", "VEsym", "VEasym" )
summarydat$covs_written <- as.character(summarydat$covs_written)
summarydat$covs_written <- factor(summarydat$covs_written, levels=unique(summarydat$covs_written))

#create table
summarydat$means_r<- round(summarydat$means, digits=2)
summarydat$lcis_r<- round(summarydat$lcis, digits=2)
summarydat$ucis_r<- round(summarydat$ucis, digits=2)

avs_bias<-summarydat[,c( "covs_written", "means_r", "lcis_r", "ucis_r", "means", "lcis", "ucis")]
avs_bias
```


# 4. Absoulte estimates figure

## a. choose variables

mod_30 is the model to use. This is the final model with all selected covariates, with age centered around 30. 

```{r absolute parameter ests all}
pars_ab<-c("FOI_p_ref", "FOI_p_cov1", "FOI_p_cov2", "FOI_p_cov3_30yrs", "FOI_p_cov4", "FOI_p_cov5", 
           "PS_p_ref", "PS_p_cov1", "PS_p_cov2", "PS_p_cov3_30yrs", "PS_p_cov4", "PS_p_cov5", 
           "VE_in_ref",  "VE_in_cov1", "VE_in_cov2", "VE_in_cov3_30yrs", "VE_in_cov4", "VE_in_cov5",
           "VE_pr_ref", "VE_pr_cov1", "VE_pr_cov2", "VE_pr_cov3_30yrs", "VE_pr_cov4", "VE_pr_cov5", 
           "VE_sym_ref", "VE_sym_cov1", "VE_sym_cov2", "VE_sym_cov3_30yrs", "VE_sym_cov4", "VE_sym_cov5", 
           "VE_asym_ref", "VE_asym_cov1", "VE_asym_cov2", "VE_asym_cov3_30yrs", "VE_asym_cov4", "VE_asym_cov5" 
           )
```

```{r convert to dataframe}
dat_mod_fin<-as.data.frame(mod_30)
dat_mod_fin<-dat_mod_fin%>%
  select(c(pars_ab))
dat_mod_fin
```

## b. create summary data for figures

Note using cov3_30yrs as this shows the difference between a 60 and 30 year old.

```{r create summary dataset}
summary_dat_means<-
  dat_mod_fin%>%
summarise(across(c(FOI_p_ref, FOI_p_cov1, FOI_p_cov2, FOI_p_cov3_30yrs, FOI_p_cov4, FOI_p_cov5,  
  PS_p_ref, PS_p_cov1, PS_p_cov2, PS_p_cov3_30yrs, PS_p_cov4, PS_p_cov5, 
  VE_in_ref, VE_in_cov1, VE_in_cov2, VE_in_cov3_30yrs, VE_in_cov4, VE_in_cov5,  
  VE_pr_ref, VE_pr_cov1, VE_pr_cov2, VE_pr_cov3_30yrs, VE_pr_cov4, VE_pr_cov5, 
  VE_sym_ref, VE_sym_cov1, VE_sym_cov2, VE_sym_cov3_30yrs, VE_sym_cov4, VE_sym_cov5,
  VE_asym_ref, VE_asym_cov1, VE_asym_cov2, VE_asym_cov3_30yrs, VE_asym_cov4, VE_asym_cov5),
             .fns = list(mean=mean))
)

summary_dat_lci<-
  dat_mod_fin%>%
summarise(across(c(FOI_p_ref, FOI_p_cov1, FOI_p_cov2, FOI_p_cov3_30yrs, FOI_p_cov4, FOI_p_cov5,  
  PS_p_ref, PS_p_cov1, PS_p_cov2, PS_p_cov3_30yrs, PS_p_cov4, PS_p_cov5, 
  VE_in_ref, VE_in_cov1, VE_in_cov2, VE_in_cov3_30yrs, VE_in_cov4, VE_in_cov5,  
  VE_pr_ref, VE_pr_cov1, VE_pr_cov2, VE_pr_cov3_30yrs, VE_pr_cov4, VE_pr_cov5, 
  VE_sym_ref, VE_sym_cov1, VE_sym_cov2, VE_sym_cov3_30yrs, VE_sym_cov4, VE_sym_cov5,
  VE_asym_ref, VE_asym_cov1, VE_asym_cov2, VE_asym_cov3_30yrs, VE_asym_cov4, VE_asym_cov5),
             .fns = list(lci=lci))
)

summary_dat_uci<-
  dat_mod_fin%>%
summarise(across(c(FOI_p_ref, FOI_p_cov1, FOI_p_cov2, FOI_p_cov3_30yrs, FOI_p_cov4, FOI_p_cov5,  
  PS_p_ref, PS_p_cov1, PS_p_cov2, PS_p_cov3_30yrs, PS_p_cov4, PS_p_cov5, 
  VE_in_ref, VE_in_cov1, VE_in_cov2, VE_in_cov3_30yrs, VE_in_cov4, VE_in_cov5,  
  VE_pr_ref, VE_pr_cov1, VE_pr_cov2, VE_pr_cov3_30yrs, VE_pr_cov4, VE_pr_cov5, 
  VE_sym_ref, VE_sym_cov1, VE_sym_cov2, VE_sym_cov3_30yrs, VE_sym_cov4, VE_sym_cov5,
  VE_asym_ref, VE_asym_cov1, VE_asym_cov2, VE_asym_cov3_30yrs, VE_asym_cov4, VE_asym_cov5),
             .fns = list(uci=uci))
)

params= c(rep("FOI", 6), rep("PS", 6), rep("VEin", 6), rep("VEpr", 6), rep("VEsym", 6), rep("VEasym", 6))
covariates=c(pars_ab)
means=as.numeric(summary_dat_means[1,])
ucis=as.numeric(summary_dat_uci[1,])
lcis=as.numeric(summary_dat_lci[1,])

summarydat<-data.frame(params, covariates, means, ucis, lcis)

summarydat$covariates <- as.character(summarydat$covariates)
summarydat$covariates <- factor(summarydat$covariates, levels=unique(summarydat$covariates))

summarydat$covs_written<-c("FOI ref.", "HCW (0 COVID-19 patients)", "HCW (1+ COVID-19 patients)", "Age (60 years)", "Obese (yes)", "Non-white (yes)",
  "Prob. symptoms ref.", "HCW (0 COVID-19 patients)", "HCW (1+ COVID-19 patients)", "Age (60 years)", "Obese (yes)", "Non-white (yes)",
  "VEin ref.", "HCW (0 COVID-19 patients)", "HCW (1+ COVID-19 patients)", "Age (60 years)", "Obese (yes)", "Non-white (yes)",
  "VEpr ref.", "HCW (0 COVID-19 patients)", "HCW (1+ COVID-19 patients)", "Age (60 years)", "Obese (yes)", "Non-white (yes)",
  "VEsym ref.", "HCW (0 COVID-19 patients)", "HCW (1+ COVID-19 patients)", "Age (60 years)", "Obese (yes)", "Non-white (yes)",
  "VEasym ref.", "HCW (0 COVID-19 patients)", "HCW (1+ COVID-19 patients)", "Age (60 years)", "Obese (yes)", "Non-white (yes)"
  )


summarydat$covs_written <- as.character(summarydat$covs_written)
summarydat$covs_written <- factor(summarydat$covs_written, levels=unique(summarydat$covs_written))
```


```{r table}
summarydat$means_r<- round(summarydat$means, digits=2)
summarydat$lcis_r<- round(summarydat$lcis, digits=2)
summarydat$ucis_r<- round(summarydat$ucis, digits=2)

summarydat[,c( "covs_written", "means_r", "lcis_r", "ucis_r")]
```

## c. create figures

```{r figure setup}
summarydat$ref="2. No"
summarydat$ref[c(1,7,13,19, 25, 31)]<-"1. Yes"
FOI_dat<-summarydat[which(summarydat$params=="FOI"),]
PS_dat<-summarydat[which(summarydat$params=="PS"),]
VEin_dat<-summarydat[which(summarydat$params=="VEin"),]
VEpr_dat<-summarydat[which(summarydat$params=="VEpr"),]
VEsym_dat<-summarydat[which(summarydat$params=="VEsym"),]
VEasym_dat<-summarydat[which(summarydat$params=="VEasym"),]

```

So the non-significant variables don't show an estimate (may mislead), convert them to NA. 

```{r remove non significant predictors}
FOI_dat[which(FOI_dat$covariates == "FOI_p_cov4" | FOI_dat$covariates == "FOI_p_cov5" ),]$means<-NA
FOI_dat[which(FOI_dat$covariates == "FOI_p_cov4" | FOI_dat$covariates == "FOI_p_cov5" ),]$lcis<-NA
FOI_dat[which(FOI_dat$covariates == "FOI_p_cov4" | FOI_dat$covariates == "FOI_p_cov5" ),]$ucis<-NA


VEin_dat[which(VEin_dat$covariates == "VE_in_cov4" | VEin_dat$covariates == "VE_in_cov5" | VEin_dat$covariates == "VE_in_cov1" | VEin_dat$covariates == "VE_in_cov2" ),]$means<-NA
VEin_dat[which(VEin_dat$covariates == "VE_in_cov4" | VEin_dat$covariates == "VE_in_cov5" | VEin_dat$covariates == "VE_in_cov1" | VEin_dat$covariates == "VE_in_cov2" ),]$lcis<-NA
VEin_dat[which(VEin_dat$covariates == "VE_in_cov4" | VEin_dat$covariates == "VE_in_cov5" | VEin_dat$covariates == "VE_in_cov1" | VEin_dat$covariates == "VE_in_cov2" ),]$ucis<-NA
```

FOI 
```{r FOI figure}
yinter=FOI_dat$means[1]
FOI_plot<-ggplot(data=FOI_dat, aes(x=covariates, y=means, ymin=lcis, ymax=ucis, color=covariates)) +
        geom_pointrange(size=1.1) + 
        coord_flip(ylim = c(0,1)) + 
        labs(y="Mean (95% CI)", x="", title="Force of Infection")+
        scale_y_continuous(breaks=seq(0,1,0.1), labels = c("0", "", "0.2", "", "0.4", "", "0.6", "", "0.8", "", "1.0"))+
        scale_x_discrete(limits=rev ,
                         labels=c(c("Non-white", "Obese", "60 year old", "HCW (1+ COVID-19 patients)", "HCW (0 COVID-19 patients)", "Reference*"))
                         )+
        theme_bw() + 
        theme(strip.placement = "null",  
              legend.position = "null",
              plot.title = element_text(hjust = 0.5, size=18),
              strip.background = element_blank(),
              strip.text = element_blank(),
              axis.text = element_text(size = 16),
              axis.title = element_text(size = 16))+
  geom_hline(yintercept = yinter, linetype="dashed", color="red", alpha=0.7, size=1)+
  scale_color_manual(
    values =c("red", "black", "black", "black", "grey", "grey"))
FOI_plot
```
PS
```{r PS figure}
yinter=PS_dat$means[1]
PS_plot<-ggplot(data=PS_dat, aes(x=covariates, y=means, ymin=lcis, ymax=ucis, color=covariates)) +
        geom_pointrange(size=1.1) + 
        coord_flip(ylim = c(0,1)) + 
        labs(y="Mean (95% CI)", x="", title="Probability of Symptoms")+
        scale_y_continuous(breaks=seq(0,1,0.1), labels = c("0", "", "0.2", "", "0.4", "", "0.6", "", "0.8", "", "1.0"))+
        scale_x_discrete(limits=rev ,
                         labels=c(c("Non-white", "Obese", "60 year old", "HCW (1+ COVID-19 patients)", "HCW (0 COVID-19 patients)", "Reference*"))
                         )+
        theme_bw() + 
        theme(strip.placement = "null",  
              legend.position = "null",
              plot.title = element_text(hjust = 0.5, size=18),
              strip.background = element_blank(),
              strip.text = element_blank(),
              axis.text = element_text(size = 16),
              axis.title = element_text(size = 16))+
  geom_hline(yintercept = yinter, linetype="dashed", color="red", alpha=0.7, size=1)+
  scale_color_manual(
    values =c("red", rep("black",5)))
PS_plot
```
VEin
```{r VEin figure}
yinter=VEin_dat$means[1]
VEin_plot<-ggplot(data=VEin_dat, aes(x=covariates, y=means, ymin=lcis, ymax=ucis, color=covariates)) +
        geom_pointrange(size=1.1) + 
        coord_flip(ylim = c(0,1)) + 
        labs(y="Mean (95% CI)", x="", title="VE against infection")+
        scale_y_continuous(breaks=seq(0,1,0.1), labels = c("0", "", "0.2", "", "0.4", "", "0.6", "", "0.8", "", "1.0"))+
        scale_x_discrete(limits=rev ,
                         labels=c(c("Non-white", "Obese", "60 year old", "HCW (1+ COVID-19 patients)", "HCW (0 COVID-19 patients)", "Reference*"))
                         )+
        theme_bw() + 
        theme(strip.placement = "null",  
              legend.position = "null",
              plot.title = element_text(hjust = 0.5, size=18),
              strip.background = element_blank(),
              strip.text = element_blank(),
              axis.text = element_text(size = 16),
              axis.title = element_text(size = 16))+
  geom_hline(yintercept = yinter, linetype="dashed", color="red", alpha=0.7, size=1)+
  scale_color_manual(
    values =c("red", "grey", "grey", "black", "grey",  "grey"))
VEin_plot
```
VEpr
```{r VEpr figure}
yinter=VEpr_dat$means[1]
VEpr_plot<-ggplot(data=VEpr_dat, aes(x=covariates, y=means, ymin=lcis, ymax=ucis, color=ref)) +
        geom_pointrange(size=1.1) + 
        coord_flip(ylim = c(0,1)) + 
        labs(y="Mean (95% CI)", x="", title="VE against progression to symptoms")+
        scale_y_continuous(breaks=seq(0,1,0.1), labels = c("0", "", "0.2", "", "0.4", "", "0.6", "", "0.8", "", "1.0"))+
        scale_x_discrete(limits=rev ,
                         labels=c(c("Non-white", "Obese", "60 year old", "HCW (1+ COVID-19 patients)", "HCW (0 COVID-19 patients)", "Reference*"))
                         )+
        theme_bw() + 
        theme(strip.placement = "null",  
              legend.position = "null",
              plot.title = element_text(hjust = 0.5, size=18),
              strip.background = element_blank(),
              strip.text = element_blank(),
              axis.text = element_text(size = 16),
              axis.title = element_text(size = 16))+
  geom_hline(yintercept = yinter, linetype="dashed", color="red", alpha=0.7, size=1)+
  scale_color_manual(
    values =c("red", rep("black",5)))
VEpr_plot
```
